### 2004-07-12 Documentation/part2.tex
Minor changes

{\tt \tb @cross\_file = XCC.dat}\\

@@

We now give a short introduction to the syntax of AEDL.
 The first eight characters have to be {\tt \%!GLoBES}
in order to avoid parsing megabytes of chunk
 and producing thousands of error messages. 
%
Comments can be used such as in C:
\begin{quote}
{\tt /* This starts a comment\\
 and here the comment ends */
}
\end{quote}
There are pre-defined variables which all start with {\tt \$}. Their range
is also checked. For example,  {\tt 
\$bins} can be only between $0$ and $150$. If one uses a {\tt float} quantity where  an {\tt int} is expected, the {\tt float} will be converted to an {\tt int} in the same way as in C.  For example, we have scalar variables
\begin{quote}
{\tt
\$bins = 10\\
\$baseline = 1200.0
}
\end{quote}
and simple lists
\begin{quote}
{\tt
\$densitytab=\{1.0,2.2343,3.3432\} 
}
\end{quote}
%
Since there are often groups of data which we want to refer to later,
environments can be used. This is illustrated 
with the channel definition part:
\begin{quote}
{\tt channel(\#ch1)<\\
\tb  $\ldots$\\
>
}
\end{quote}
The first part is the type of environment, which is {\tt channel} here. 
There are the following types of environment in AEDL:
\begin{quote}
{\tt flux\\
cross\\
channel\\
energy\\
rule
}
\end{quote}
Besides the environment type, there is a user-defined name 
beginning with {\tt \#}
in the above example: {\tt \#ch1}. It can be used later to refer to the 
channel defined in {\tt <$\ldots$>}. Those names are so-called 
``automatic variables'' and have to start with {\tt \#}. Note that these names have to be unique and can only be refered to after their definition.
However, similar to C, one can give a header definition before:
\begin{quote}
{\tt    channel(\#ch2)<>}
\end{quote}
Now one can refer to the name {\tt \#ch2}, while the actual channel definition comes later. The internal representation of this automatic
variable is a number, which obtains its value from a counter for each type of environment. For example, for {\tt channel} the counter is {\tt numofchannels}. The counter keeps track of how many different names 
there are for one type of environment, which means that it counts the number of channels, rules, energy resolution functions \etc . Thus, the automatic
variables are numbered in the order of their definition, and the number
can later be used to refer to them in the C code (from $0$ to {\tt numof...}$-1$). Within each environment type, there are several 
variables beginning with {\tt @}, which can only be used within the 
appropriate type of environment. In many cases, 
they have a special syntax, such as {\tt @channel}.

@@

As another feature of AEDL one can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
 for nested includes, too. The included file is not required to begin 
 with {\tt \%!GLoBES} to facilitate cut and paste:
\begin{quote}
{\tt include ./file\_1}
\end{quote}
With this include mechanism, one can use constructions such as 
\begin{quote}
{\tt    include NuFact.gls\\
        \#NEXT\#\\
        include JHFHK.gls
}
\end{quote}
in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. 
Even if one uses the
automatic variable {\tt \#CC} in both experiments, 
but the cross section data are different (for example, because of different target nuclei), the correct 
cross section data will be applied to each of the experiments. 
Note that, alternatively, one can 
also load both files successively.

Furthermore, one can define constants such as
\begin{quote}
{\tt
Pi = 3.141
}
\end{quote}
and use some simple algebraic manipulations such as
\begin{quote}
{\tt
Pi+1\\
\verb+Pi^2+\\
sin(Pi/2)\\
}
\end{quote}
The following mathematical functions from {\tt <math.h>} are available: 
{\tt sin}, {\tt cos}, {\tt atan}, {\tt ln}, {\tt exp}, {\tt sqrt}. 
These functions can be used everywhere, where
otherwise only a scalar number would appear. However, they can not be
applied to lists, such as {\tt sin(\{1,2,3\})} will not work.

@@

For a built-in neutrino source, one has to specify which
built-in spectrum has to be used, as well as its parameters. The software
will then automatically calculate the neutrino spectrum. Note that in this
case, there is no degree of freedom in the choice of the source units.
Currently, two built-in fluxes are available: $\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}). In these cases, the muon energy (enery of the parent particle) has to be specified together with the number of useful decays  per year. Thus, an example to set up a neutrino factory flux is
\begin{quote}
{\tt flux(\#mu\_plus)<\\
\tb  @builtin = 1\\
\tb  @parent\_energy = 50.0\\
\tb  @stored\_muons = 5.33e+20\\
\tb  @time = 8.0\\
>}
\end{quote}
%
For a user-defined flux, one has to give it the file name:
\begin{quote}
{\tt flux(\#user)<}\\
{\tt \tb @flux\_file = user\_file\_1.dat\\
\tb @time = 2.0\\
\tb @power = 4.0\\
\tb @norm = 1e+8}\\
{\tt >}
\end{quote}
In this case, the {\tt @norm} variable is an overall normalization which defines a conversion factor from the fluxes in the file to the units in \GLOBES . In general, there are many ways to give the source power of a 
neutrino source, such as neutrinos per proton on target per area per time frame. Right now, each flux has its own normalization factor, which is
not always straightfoward to calculate. Often, one has to take into account
many things, such as the number of target particles per unit mass. 
In addition, the fluxes will be rescaled by $1/L^2$, which means that the
normalization must contain a factor $L_0^2$. Here $L_0$ is the distance from the source for which the flux is given to the actual neutrino production region. At the end, it is left to the user to ensure that the 
numbers in the flux file give, after the multiplication with {\tt @norm}, 
the proper numbers of produced neutrinos corresponding to the chosen target power {\tt @power}. Usually this adjustment of {\tt @norm} is performed by comparison with known energy spectra for a specific experiment.

@@

Cross sections\index{Cross sections} will later be used as part of the 
channel definition (see \Sec~\ref{sec:channel}). Similar to the source 
fluxes, they are provided by the user as a data file:
\begin{quote}
{\tt cross(\#name)<}\\
{\tt \tb @cross\_file =user\_file\_1.dat}\\
{\tt >}
\end{quote}  
This cross section can later be refered to by {\tt \#name}.

@@

A simple channel definition in \GLOBES\ consists of the flux,
the CP-sign of the initial state, the initial flavor, the final flavor,
the cross sections, and the energy resolution function. In order to refer to
flux, cross sections, and energy resolution functions, they have to be 
defined before with their {\tt \#name} in the respective environments. 
Thus, a simple definition of a channel is
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
>}
\end{quote}
Note that the energy environment will be described in the next section. 
In addition, one can define pre- and post-smearing effects together
with the channels, which will also be introduced together with the
energy resolution function in the next section.

@@

Eventually, a rule looks like
\begin{quote}
{\tt rule(\#rule\_1)<\\
\tb @signal = 0.5 @ \#channel\_1\\
\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3\\
\tb @signalerror =       0.001  :       0.01\\
\tb @backgroundcenter =  0.1 :       0.0\\
\tb @backgrounderror =   0.001 :       0.01\\
\tb @errordim = 1\\
\tb @energy\_window = 4.0 : 50.0\\ 
>}
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{User-defined systematics}
%\label{sec:ud_error_dim}.

>>>>

{\tt \tb @cross\_file = "XCC.dat"}\\

@@

We now give a short introduction to the syntax of AEDL.
 The first eight characters have to be {\tt \%!GLoBES}
in order to avoid parsing megabytes of chunk
 and producing thousands of error messages. 
%
Comments can be used such as in C:
\begin{quote}
{\tt /* This starts a comment\\
 and here the comment ends */
}
\end{quote}
There are pre-defined variables which all start with {\tt \$}. Their range
is also checked. For example,  {\tt 
\$bins} can be only between $0$ and $500$~\footnote{The upper limit is 
there only for safety reasons, the memory is allocated dynamically.}. If one uses a {\tt float} quantity where  an {\tt int} is expected, the {\tt float} will be converted to an {\tt int} in the same way as in C.  For example, we have scalar variables
\begin{quote}
{\tt
\$bins = 10\\
\$baseline = 1200.0
}
\end{quote}
and simple lists
\begin{quote}
{\tt
\$densitytab=\{1.0,2.2343,3.3432\} 
}
\end{quote}
%
Since there are often groups of data which we want to refer to later,
environments can be used. This is illustrated 
with the channel definition part:
\begin{quote}
{\tt channel(\#ch1)<\\
\tb  $\ldots$\\
>
}
\end{quote}
The first part is the type of environment, which is {\tt channel} here. 
There are the following types of environment in AEDL:
\begin{quote}
{\tt flux\\
cross\\
channel\\
energy\\
rule
}
\end{quote}
Besides the environment type, there is a user-defined name 
beginning with {\tt \#}
in the above example: {\tt \#ch1}. It can be used later to refer to the 
channel defined in {\tt <$\ldots$>}. Those names are so-called 
``automatic variables'' and have to start with {\tt \#}. Note that these names have to be unique and can only be refered to after their definition.
However, similar to C, one can give a header definition before:
\begin{quote}
{\tt    channel(\#ch2)<>}
\end{quote}
Now one can refer to the name {\tt \#ch2}, while the actual channel definition comes later. The internal representation of this automatic
variable is a number, which obtains its value from a counter for each type of environment. For example, for {\tt channel} the counter is {\tt numofchannels}. The counter keeps track of how many different names 
there are for one type of environment, which means that it counts the number of channels, rules, energy resolution functions \etc . Thus, the automatic
variables are numbered in the order of their definition, and the number
can later be used to refer to them in the C code (from $0$ to {\tt numof...}$-1$). Within each environment type, there are several 
variables beginning with {\tt @}, which can only be used within the 
appropriate type of environment. In many cases, 
they have a special syntax, such as {\tt @channel}.

@@

As another feature of AEDL one can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
 for nested includes, too. The included file is not required to begin 
 with {\tt \%!GLoBES} to facilitate cut and paste:
\begin{quote}
{\tt include "./file\_1"}
\end{quote}
With this include mechanism, one can use constructions such as 
\begin{quote}
{\tt    include  "NuFact.gls"\\
        \#NEXT\#\\
        include   "JHFHK.gls"
}
\end{quote}
in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. As one sees from the example above
file names are always quoted in \AEDL.\index{File names}
Even if one uses the
automatic variable {\tt \#CC} in both experiments, 
but the cross section data are different (for example, because of different target nuclei), the correct 
cross section data will be applied to each of the experiments. 
Note that, alternatively, one can 
also load both files successively.

Furthermore, one can define constants such as
\begin{quote}
{\tt
Pi = 3.141
}
\end{quote}
These constants can not only be defined within one \AEDL\ file but also
by the calling C program, which allows to use a simple but powerful variable
substitution mechanism as described in~\Sec~\ref{sec:aedlparams}.
\index{External \AEDL\ parameters}

Alos some simple algebraic manipulations are possible, such as
\begin{quote}
{\tt
Pi+1\\
\verb+Pi^2+\\
sin(Pi/2)\\
}
\end{quote}
The following mathematical functions from {\tt <math.h>} are available: 
{\tt sin}, {\tt cos}, {\tt tan}, {\tt asin}, {\tt acos}, {\tt atan}, 
{\tt log}, {\tt log10}, {\tt exp}, {\tt sqrt}. 
These functions can be used everywhere, where
otherwise only a scalar number would appear. However, they can not be
applied to lists, such as {\tt sin(\{1,2,3\})} will not work.

@@

For a built-in neutrino source, one has to specify which
built-in spectrum has to be used, as well as its parameters. The software
will then automatically calculate the neutrino spectrum. Note that in this
case, there is no degree of freedom in the choice of the source units.
Currently, two built-in fluxes are available: $\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}). In these cases, the muon energy (enery of the parent particle) has to be specified together with the number of useful decays  per year. Thus, an example to set up a neutrino factory flux is
\begin{quote}
{\tt flux(\#mu\_plus)<\\
\tb  @builtin = 1\\
\tb  @parent\_energy = 50.0\\
\tb  @stored\_muons = 5.33e+20\\
\tb  @time = 8.0\\
>}
\end{quote}
%
For a user-defined flux, one has to give it the file name:
\begin{quote}
{\tt flux(\#user)<}\\
{\tt \tb @flux\_file = "user\_file\_1.dat"\\
\tb @time = 2.0\\
\tb @power = 4.0\\
\tb @norm = 1e+8}\\
{\tt >}
\end{quote}
In this case, the {\tt @norm} variable is an overall normalization which defines a conversion factor from the fluxes in the file to the units in \GLOBES . In general, there are many ways to give the source power of a 
neutrino source, such as neutrinos per proton on target per area per time frame. Right now, each flux has its own normalization factor, which is
not always straightfoward to calculate. Often, one has to take into account
many things, such as the number of target particles per unit mass. 
In addition, the fluxes will be rescaled by $1/L^2$, which means that the
normalization must contain a factor $L_0^2$. Here $L_0$ is the distance from the source for which the flux is given to the actual neutrino production region. At the end, it is left to the user to ensure that the 
numbers in the flux file give, after the multiplication with {\tt @norm}, 
the proper numbers of produced neutrinos corresponding to the chosen target power {\tt @power}. Usually this adjustment of {\tt @norm} is performed by comparison with known energy spectra for a specific experiment.

@@

Cross sections\index{Cross sections} will later be used as part of the 
channel definition (see \Sec~\ref{sec:channel}). Similar to the source 
fluxes, they are provided by the user as a data file:
\begin{quote}
{\tt cross(\#name)<}\\
{\tt \tb @cross\_file ="user\_file\_1.dat"}\\
{\tt >}
\end{quote}  
This cross section can later be refered to by {\tt \#name}.

@@

A simple channel definition in \GLOBES\ consists of the flux,
the CP-sign of the initial state, the initial flavor, the final flavor,
the cross sections, and the energy resolution function. In order to refer to
flux, cross sections, and energy resolution functions, they have to be 
defined before with their {\tt \#name} in the respective environments. 
Thus, a simple definition of a channel is
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
>}
\end{quote}

\index{Oscillation!switching off}
It is also possible to define a channel as no-oscillation by using the
prefix \GLB{NOSC\_} in either the initial  flavour or 
the final flavour, like this
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: NOSC\_muon: muon: \#cross: \#energy\\
>}
\end{quote}

 In that case
all diagonal probabilities $P_{\alpha\alpha}$ are unity and all off-diagonal
probablities  $P_{\alpha\beta}$ are zero. This is \eg\ especially useful for
neutral current events since they do not depend on any oscillation 
parameters\footnote{at least in the absence of sterile neutrinos}. The channels
marked as \GLB{NOSC\_} are only computed by a call to \GLB{glbSetRates}
and not recomputed in the subsequent fit (which calls \GLB{glbSetNewRates}).
Therefore this feature can be used to speed up the rate 
computation considerably, especially in cases where a large set
of channels exist which are only used for the computation of backgrounds.
Usually it is an excellent approximation to treat backgrounds as if they
would not be affected by oscillations.

Note that the energy environment will be described in the next section. 
In addition, one can define pre- and post-smearing effects together
with the channels, which will also be introduced together with the
energy resolution function in the next section.

@@

Eventually, a rule looks like
\begin{quote}
{\tt rule(\#rule\_1)<\\
\tb @signal = 0.5 @ \#channel\_1\\
\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3\\
\tb @signalerror =       0.001  :       0.01\\
\tb @backgroundcenter =  0.1 :       0.0\\
\tb @backgrounderror =   0.001 :       0.01\\
\tb @errordim\_sys\_off = 2\\
\tb @errordim\_sys\_on = 0\\
\tb @energy\_window = 4.0 : 50.0\\ 
>}
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Versioning of \AEDL\ files}
\label{sec:aedl_versioning}
\index{Version control}

In order to provide a handle on problems which may arise due to different
versions of \GLOBES\ and \AEDL\ files, it is possible to stamp each 
\AEDL\ file with the corresponding version number of the \GLOBES\ package with
\GLB{\$version} by
\begin{quote}
{\tt \$version="1.8.1"}
\end{quote}
This information can be accesed by the versioning functions as described in~\Sec~\ref{sec:versioning}.
